---
title: "Zillow US Mapping"
author: "Jerry Qian"
date: "April 5, 2019"
output: html_document
---

```{r setup, include=FALSE}
setwd("D:/Documents/Semester4/HONR238V/HONR238V/")
```

# Raw cleaned state housing price data
```{r}
library(dplyr)
library(tidyr)
library(lubridate)

stateHomePriceTimeSeries <- read.csv("D:/Documents/Semester4/HONR238V/HONR238V/Data/Home_Values/State_Zhvi_AllHomes.csv") %>% 
  select(-c("RegionID", "SizeRank")) %>%
  mutate(recessionStart = X2007.11, recessionEnd = X2009.06, recessionDiff = X2009.06-X2007.11) %>%
  gather(date, price, -c(fips,RegionName,recessionStart,recessionEnd,recessionDiff)) %>%
  mutate(date=as.Date(paste(date,".01",sep=''), "X%Y.%m.%d"), year = as.integer(year(date)))
  
head(stateHomePriceTimeSeries, 100)
```

# Raw cleaned state population data
```{r}
library(dplyr)
library(tidyr)
library(lubridate)
library(cdlTools)

stateLandArea <- read.csv("D:/Documents/Semester4/HONR238V/HONR238V/Data/state_land_area.csv") %>%
  mutate(squareMiles = as.numeric(sub(",", "", as.character(squareMiles), fixed = TRUE)), fips = as.integer(fips(state)))

head(stateLandArea, 1000)

statePopulations <- read.csv("D:/Documents/Semester4/HONR238V/HONR238V/Data/state_populations.csv") %>%
  mutate(date = as.Date(DATE)) %>%
  select(-c(DATE)) %>%
  filter(year(date) >= 1996) %>%
  gather(state, population, -c(date)) %>%
  mutate(year = as.integer(year(date)), state = substr(state, 0, 2), fips = as.integer(fips(state)), population = population*1000) %>%
  select(-c(state)) %>%
  left_join(stateLandArea, by=c("fips")) %>%
  mutate(density = population/squareMiles)

statePopulationsPrices <- statePopulations %>%
  select(-c(year)) %>%
  left_join(stateHomePriceTimeSeries_standardized, by=c("fips", "date"))

head(statePopulationsPrices, 1000)

#State_Zri_AllHomesPlusMultifamily.csv

```


#--------------------------------------------------------------------------------#

## Static map at chosen date
```{r}
library(gganimate)
library(usmap)
library(ggplot2)

#stateHomePriceTimeSeries2000 <- stateHomePriceTimeSeries %>% filter(date == as.Date("2000-04-01"))

#plot_usmap(data = stateHomePriceTimeSeries2000, values = "price", lines = "black") + 
  #scale_fill_continuous(low = "white", high = "red", name = "Home Price", label = scales::comma) + 
  #theme(legend.position = "right") + ggtitle("US Home Prices by State")

#statePopulations2000 <- statePopulations %>% filter(year(date) == 2017)
#plot_usmap(data = statePopulations2000, values = "density", lines = "black") + 
  #scale_fill_continuous(low = "white", high = "blue", name = "Population Density", label = scales::comma) + 
  #theme(legend.position = "right", plot.title = element_text(hjust = 0.5)) + ggtitle("Population Density by State") 


plot_usmap(data = stateHomePriceTimeSeries_standardized, values = "percentDiff", lines = "black") + 
  scale_fill_gradient2(low = "red", mid="white", high = "green", midpoint=0, name = "Percent Difference") + 
  theme(legend.position = "right", plot.title = element_text(hjust = 0.5)) + ggtitle("Home Price Percent Difference from Recession by State") 

```


## Animated map raw prices
```{r}

library(ggplot2)
library(gganimate)
library(usmap)

plot_usmap(data = stateHomePriceTimeSeries, values = "price", lines = "black") + 
  scale_fill_continuous(low = "white", high = "red", name = "Home Price", label = scales::comma) + 
  theme(legend.position = "right") + 
  labs(title = 'Date: {frame_time}    Average US Home Prices by State') + 
  transition_time(date)


```


# Standardization
```{r}
library(dplyr)
library(lubridate)


yearMean_stateHomePriceTimeSeries <- stateHomePriceTimeSeries %>%
  group_by(date) %>% 
  mutate(meanPrice = mean(price, na.rm = TRUE), std = sd(price, na.rm = TRUE), meanDiff = mean(recessionDiff, na.rm = TRUE), stdDiff = sd(recessionDiff, na.rm = TRUE)) %>% select(fips,date,meanPrice,std, meanDiff,stdDiff)

stateHomePriceTimeSeries_standardized <- stateHomePriceTimeSeries %>% 
  left_join(yearMean_stateHomePriceTimeSeries, by=c("fips", "date")) %>%
  mutate(standardizedPrice = (price - meanPrice)/std, percentDiff = (recessionDiff)/recessionStart, year = year(date))

head(stateHomePriceTimeSeries_standardized, 55)


```

## Line graph standardized prices per year
```{r}
library(tidyverse)
library(ggplot2)
library(ggrepel)




ggplot(stateHomePriceTimeSeries_standardized, aes(x=date, y=standardizedPrice, color=RegionName)) + 
  geom_line() + 
  theme(legend.position="none", plot.title = element_text(hjust = 0.5)) + 
  ggtitle("Standardized Average Home Prices by State") + ylab("standardized price") + 
  geom_text(aes(label = RegionName), data = stateHomePriceTimeSeries_standardized %>% filter(abs(standardizedPrice) > .5) %>% filter(as.character(date)=="2019-02-01"), check_overlap = TRUE)

falling <- stateHomePriceTimeSeries_standardized %>% filter(percentDiff < -.07)
ggplot(falling , aes(x=date, y=price, color=RegionName)) + 
  geom_line() + 
  theme(legend.position="right", plot.title = element_text(hjust = 0)) + 
  ggtitle("Average Home Prices by States On Decline during Great Recession") + ylab("average price") #+ 
  #geom_text_repel(aes(label = RegionName), data = falling %>% filter(as.character(date)=="2019-02-01"), check_overlap = FALSE, hjust="outward")

rising <- stateHomePriceTimeSeries_standardized %>% filter(percentDiff > .005)
ggplot(rising, aes(x=date, y=price, color=RegionName)) + 
  geom_line() + 
  theme(legend.position="none", plot.title = element_text(hjust = 0.8)) + 
  ggtitle("Average Home Prices by States Growing during Great Recession") + ylab("average price") + 
  geom_text(aes(label = RegionName), data = rising %>% filter(as.character(date)=="2019-02-01"), check_overlap = TRUE)



```

## Animated map standardized prices per year
```{r}

library(ggplot2)
library(gganimate)
library(usmap)

plot_usmap(data = stateHomePriceTimeSeries_standardized, values = "standardizedPrice", lines = "black") + 
  scale_fill_continuous(low = "white", high = "red", name = "Z-Score") + 
  theme(legend.position = "right") + 
  labs(title = 'Date: {frame_time}   Standardized (relative) Average US Home Prices by State') + 
  transition_time(date)


```

# Standardization of Population
```{r}
library(dplyr)
library(lubridate)


yearMean_statePopulations <- statePopulations %>%
  dplyr::group_by(date) %>% mutate(meanPopulation = mean(population, na.rm = TRUE), std = sd(population, na.rm = TRUE)) %>% select(fips,date,meanPopulation,std)

statePopulations_standardized <- statePopulations %>% 
  left_join(yearMean_statePopulations, by=c("fips", "date")) %>%
  mutate(standardizedPopulation = (population - meanPopulation)/std)

head(statePopulations_standardized, 55)


```

## Animated map population per year
```{r}

library(ggplot2)
library(gganimate)
library(usmap)

plot_usmap(data = statePopulations, values = "density", lines = "black") + 
  scale_fill_continuous(low = "white", high = "blue", name = "Residents/square mile") + 
  theme(legend.position = "right") + 
  labs(title = 'Date: {frame_time}   Population Density by State') + 
  transition_time(date)


```


## Animated scatterplot price vs population density
```{r}
library(ggplot2)
library(gganimate)
#library(usmap)

stateHomePricePopulationDensity <- stateHomePriceTimeSeries %>%
  left_join(statePopulations %>% select(-c(date)), by=c("fips", "year"))
head(stateHomePricePopulationDensity)
head(statePopulationsPrices)
#broken
ggplot(statePopulationsPrices, aes(x=density, y=standardizedPrice, color=RegionName)) + geom_point() + geom_smooth(method=lm) +
   labs(title = 'Date: {frame_time}   Population Density vs Average Home Price by State') + transition_time(date)

```


#test animated table

```{r}
library(ggplot2)
library(gganimate)
library(usmap)

ggplot(stateHomePriceTimeSeries, aes(x=fips, y=price)) + geom_point() + 
   labs(title = 'Year: {frame_time}') + transition_time(date)

```






